\bf\mftitle

### 2011-03-01 doc/jInfer/twostep/twostep.tex
Fixed some sentences from previous version.

%#### INTRO ############################################
\section{Introduction}
In \jmodule{TwoStepSimplifier} we provide user a complex solution to grammar simplification problem.
While this is the only simplifier bundled with jInfer, its strong modularity ensures that it can be extended easily. For most of the points of extension we indeed provide multiple implementations. These vary from trivial proof-of-concept to for example $k,h$-context method for grammar rules inference (see \cite{ahonen}). Strong emphasis is put on generic automata, therefore it is easy to plug-in any automaton merging state algorithm.
In this document, we will walk over all modules and submodules in simplifier and describe them briefly.

%#### FACTORY ############################################
\section{Factory pattern usage} \label{section_factory}
Lets stay a while at factory pattern usage.
In \jmodule{TwoStepSimplifier} we employ factory pattern (as described in \cite[section 4.2]{archdoc}) to divide module into submodules.
Since service providing classes are being kept as singletons in the NB platform, we use them as factories:
For example:
\begin{verbatim}
@ServiceProvider(service = AutomatonSimplifierFactory.class)
public interface AutomatonSimplifierFactory extends 
                     NamedModule, Capabilities, UserModuleDescription {
  <T> AutomatonSimplifier<T> create();
}
\end{verbatim}
The real submodule interface (called familiarly \emph{worker} interface) is the \code{AutomatonSimplifier<T>} interface (of which instance is returned by factory).
For now, it is not important how the worker interface looks like, lets just examine the factory.

In \jmodule{Two\.Step\.Sim\.pli\.fier} we design service providing factory interfaces so that they extend \code{Na\.med\.Mo\.du\.le}, \code{Ca\.pa\.bi\.li\.ties} and \code{UserModuleDescription}.
That means, modules must implement methods:
\begin{verbatim}
String getName();
String getDisplayName();
String getModuleDescription();
List<String> getCapabilities();
String getUserModuleDescription();
\end{verbatim}
Nothing non-standard apart from \code{getUserModuleDescription}.
It returns user-friendly description of the module which is then displayed in properties panels.

Under certain circumstances it is useful to declare method \code{create()} generic.
The \jmodule{AutomatonSimplifier} works with \code{Automaton} instances, which are generic.
But automaton simplification does not depend on type of symbol of automaton transitions, so interface \code{AutomatonSimplifier<T>}
generic too.
Factory interface deals with this by defining the \code{create()} method generic.

@@

Lets look at \code{AutomatonMergingStateFactory}. This is factory of module which has \jmodule{AutomatonSimplifier} as a submodule.
Its create method (shortened): 
\begin{verbatim}
@Override
public ClusterProcessor<AbstractStructuralNode> create() {
  LOG.debug("Creating new ClusterProcessorAutomatonMergingState.");
  return new AutomatonMergingState(getAutomatonSimplifierFactory(),
                                   getRegexpAutomatonSimplifierFactory());
}
\end{verbatim}
Methods \code{getAutomatonSimplifierFactory} and \code{getRegexpAutomatonSimplifierFactory} are analogical. Here is the former:
\begin{verbatim}
private AutomatonSimplifierFactory getAutomatonSimplifierFactory() {
  Properties p = RunningProject.getActiveProjectProps(getName());

  return ModuleSelectionHelper.lookupImpl(AutomatonSimplifierFactory.class,
            p.getProperty(PROPERTIES_AUTOMATON_SIMPLIFIER,
                          PROPERTIES_AUTOMATON_SIMPLIFIER_DEFAULT));
}
\end{verbatim}
Cluster processor \code{AutomatonMergingState} then receives factories of \jmodule{AutomatonSimplifier} and \jmodule{Reg\.exp\.Au\.to\.ma\.ton\.Simplifier} submodules in its constructor.
Cluster processor then may create as many instances of submodule classes as it needs (maybe simplifying more than one automaton).
Thorough this document, we will mention only worker interface when describing submodule, since all factory interfaces are designed the same way as was just described.

%#### STRUCTURE ############################################
\section{Structure}
\jmodule{TwoStepSimplifier} is implemented in package \code{cz.cuni.mff.ksi.jinfer.twostep}.
Module is divided into two classes: \code{TwoStepSimplifier} (main logic) and \code{TwoStepSimplifierFactory} (lookups, interface to other modules).
The latter is registered as service provider (and implements) of \code{Simplifier} interface (defined in package \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference}).

Its main method is \code{start()} which receives Initial Grammar in form of
\begin{verbatim}
List<Element> grammar
\end{verbatim}
Each grammar rule is represented as class \code{Element}, where element is left side of rule and its \code{getSubnodes()} method
returns regular expression representing right side of rule (as described in \cite[3.1.1]{archdoc}, they are all concatenations).
Second parameter is 
\begin{verbatim}
SimplifierCallback callback
\end{verbatim}
Callback which should be invoked when the simplification is done.
On output, simplifier provides Simplified grammar as a parameter of callback function:
\begin{verbatim}
void finished(List<Element> grammar);
\end{verbatim}
Thus simplifier receives Initial Grammar and returns Simplified Grammar, both represented same way - list of elements.

\code{TwoStepSimplifierFactory.start()} creates new \code{TwoStepSimplifier} class instance, providing three factories of submodules to its constructor.
Then it calls \code{simplify(initialGrammar)} method of this instance and its result is passed as parameter to callback function.
That's all, the magic is in \code{TwoStepSimplifierFactory} class.

\section{Modular design}
\jmodule{TwoStepSimplifier} is inspired by \cite{1802522} design. Inference proceeds in two steps:
\begin{enumerate}
	\item Clustering the element instances into clusters of (probably) same elements.
	\item Inferring regular expression for each element from examples of element contents taken from all elements in a cluster.
\end{enumerate}
Clustering is delegated to \jmodule{Clusterer} submodule, and task of regular expression inference for each cluster is delegated to \jmodule{ClusterProcessor} submodule. We will examine both of them.
There is also third submodule called \jmodule{RegularExpressionCleaner} actually, its purpose is just to beautify
 output regular expressions, no inference logic is implemented there. 
Modules are drawn on fig. \ref{twostep_modules}.
\begin{figure}[tpb]
	\centering\includegraphics[scale=\myscale]{twostep_modules}
	\caption{Submodules of TwoStep simplifier with their implementations} \label{twostep_modules}
\end{figure}
We provide one \jmodule{Clusterer}, 4 \jmodule{ClusterProcessor} and 4 \jmodule{RegularExpressionCleaner} implementations.
Each of those will be explained further in this document.

@@

return simplifiedGrammar;
\end{verbatim}
First, it creates a clusterer, gives it all the rules and performs the clustering.
Then, empty list of elements is created as the simplified grammar.
For each input rule in \code{initialGrammar}, submodule \jmodule{ClusterProcessor} is invoked to infer the regular expression of that element.
Finally, regular expression cleaning is done in a submodule and new \code{Element} instance is created as a copy of processed node, but with inferred and cleaned regexp.

\paragraph{Sentinel processing}
There are some specialities in processing sentinel elements (see \cite[3.2]{archdoc}).
Sentinels can be only on right side of rules and they have to be sent to clusterer, as someone further in the chain can ask for cluster of sentinel element - which appears on right side.
But sentinels have no content, they must not be taken into account when inferring attribute properties (whether it is required or optional) - since sentinels have no attributes at all.

Now we will examine submodules for clustering, processing and cleaning.

%#### STRUCTURE->Clusterer ############################################
\subsection{Clusterer submodule}
Clustering is implemented in \code{cz.cuni.mff.ksi.jinfer.twostep.clustering} package.
One cluster is represented by the \code{Cluster<T>} class with \code{T} as type parameter of clustered items.
This class simply holds a \code{Set} of members of this cluster and one of the references as \code{representant} member.

Any clusterer has to comply to \code{Clusterer} interface.
Its purpose is to cluster bunch of elements (rules) on input, into bunch of \code{Cluster} class instances (clusters) on output.
It has methods \code{add()} and \code{addAll()} for adding items for clustering.
Main part is the method \code{cluster()}, which does the clustering itself.

As this may be a time-consuming operation, method throws \code{InterruptedException}.
Implementation should take care of checking whether thread is user interrupted (see \cite[p. 12]{archdoc}).
To be able to respond to calls to \code{T getRepresentantForItem(T item)} method, every clusterer implementation has to keep all its clusters even after the clustering is finished. Using this method, given an item the correct cluster representative for this item has to be returned.

If no such cluster exists (item was not added for clustering before), we recommend throwing an exception rather than returning \code{null}.
Missing an item probably indicates error in algorithm rather than normal workflow.
One can get all the clusters (\code{List<Cluster<T>{}>}) from clusterer by calling \code{getClusters()} method.
Basic usage of the clusterer is:
\begin{verbatim}
Clusterer <T> c = new MyContextClusterer<T>();
c.addAll(initialGrammar);
c.cluster();
...
c.getClusters();
or
c.getRepresentantForItem(x);
\end{verbatim}

%#### STRUCTURE->Clusterer->WithAttributes ############################################
\subsubsection{ClustererWithAttributes extended interface}
You might have noticed that the \jmodule{Clusterer} interface and \code{Cluster} class are generic.
They may be used as design pattern not only for clustering elements in inference process.
To address clustering of elements in more detail, we created \code{Cluster\.er\.With\.At\.tri\.bu\.tes<T, S>} interface, which extends \code{Clusterer<T>} interface.
It adds method \code{List\.<Cluster<S>{}> get\.At\.tri\.bu\.teClusters(T representant)}, implying that each representative of
type \code{T} (that is representative of some main cluster) has some ``attribute'' clusters associated with it.
Attribute clusters are of type \code{S} and can be retrieved by calling \code{getAttributeClusters(x)}.

@@

%#### STRUCTURE->ClusterProcessor ############################################
\subsection{ClusterProcessor module}
\jmodule{ClusterProcessor} takes rules of one cluster of element and somehow obtains regular expression representing that set of elements. It returns one rule - element with name set to desired name of element in schema (not all elements in cluster have to have same name, if advanced clustering scheme is used, then processor has to choose the right name for the resulting element) and with subnodes set to regular expression inferred.
It processes attributes of all elements in cluster to obtain meaningful schema attribute specification and these attributes have to be attached to the resulting element.

Interface of submodule is defined as follows.
\begin{verbatim}
public interface ClusterProcessor<T> {
  T processCluster(
          Clusterer<T> clusterer,
          List<T> rules
          ) throws InterruptedException;
}
\end{verbatim}
Why is the cluster processor given a clusterer instance?
Rules themselves contain information about which elements to process, but clusterer has more information about the topic.
Clusterer can tell you the representative for any element in the whole input (not only those elements in rules, but also those that may be on right side of rules). Also, clusterer (if it is with attributes) has information about attributes in each cluster.

@@

%#### STRUCTURE->RegularExpressionCleaner ############################################
\subsection{RegularExpressionCleaner module}
Finally we examine the \jmodule{RegularExpressionCleaner} interface.
Purpose of this submodule is only to make output regular expression corrections, to make them nicer.
For example it is common that converting automaton to regular expression by state elimination produces nested concatenations such as $(name, (person, id))$.
To convert such expression into $(name, person, id)$, one can implement this interface and connect it to work in chain.
Interface definition is straightforward here.
\begin{verbatim}
public interface RegularExpressionCleaner<T> {
  Regexp<T> cleanRegularExpression(Regexp<T> regexp);
}
\end{verbatim}
Given regular expression, return regular expression.
Conversion of regular expression is commonly a recursive task and all our implementations work this way.

@@

%#### DATA FLOW ############################################
\section{Data flow}
\begin{figure}[tpb]
	\centering\includegraphics[scale=1]{twostep_flow}
	\caption{TwoStep data flow} \label{twostep_flow}
\end{figure}
Process illustrated in fig. \ref{twostep_flow} is as follows.
\begin{enumerate}
	\item Initial Grammar on input, which is sent into \jmodule{Clusterer}.
	\item Each cluster is sent to submodule \jmodule{ClusterProcessor}, which returns regular expression for that cluster.
	\item Regular expression is sent to \jmodule{RegularExpressionCleaner} submodule for cleaning.
	\item Regular expressions representing all the processed clusters are added to the list of simplified grammar rules.
	\item List of simplified rules is returned (Simplified grammar).
\end{enumerate}

%#### SUBMODULES ############################################
\section{Submodule implementations}
%#### SUBMODULES->Iname ############################################
\subsection{Iname (ClusteterWithAttributes)} \label{section_Iname}
We cluster elements in \code{Iname\.<Abstract\.Structural\.Node, Attri\.bu\.te>} class, which takes \code{Abstract\.Structural\.No\.de} classes to cluster as main, and \code{Attribute} classes as attributes.
Clustering is done based on element's \code{getName()} equality (ignoring case).
For each rule (element), it is first clustered by finding the cluster which representative has the same name - or create new cluster with this element as representative.
Then we process right side of this element rule (that are nodes from \code{getSubnodes()}).
Since right side of rule is always concatenation (\jmodule{TwoStep} accepts only simple grammar representation), we simply take \code{.getSubnodes().getTokens()} list and iterate through it.
Each node on right side, is examined in the following way.
\begin{itemize}
  \item If it is \textbf{simple data}, throw it to \code{SimpleDataClusterer} class which is associated with main element cluster. It does nothing more, than holds all \code{SimpleData} instances in one cluster. But in future it may be replaced to cluster simple data somehow, to obtain meaningful content models in schemas.
  \item If it is \textbf{element} and it is tagged as sentinel - search main clusters to find cluster with representative of same \code{getName()}, or create new cluster with this sentinel as representative. From IGG, sentinels may be only on right sides of rules and since each element in schemas has to be defined, there must exist another element with same name, which is not sentinel (and maybe it will come to process in future). So there can't be cluster with only one sentinel element in it. 
  \item Do nothing otherwise, since it is element, it has to have its subnodes defined, and therefore it is element that is proper grammar rule and therefore it has to be somewhere in Initial Grammar, thus it is already processed or is on schedule.
\end{itemize}
Attributes of element are processed through a helper attribute clusterer, which is created for each element cluster.
We have bunch of elements of same name in a cluster and one attribute clusterer associated with this bunch.
This attribute clusterer is given all attributes instances encountered in all elements that are in bunch.
It clusters them by name, case insensitive.
Each main cluster also has \code{SimpleData} clusterer associated with it.
Whole scheme is on fig. \ref{clusterer_structure}.
\begin{figure}[tpb]
	\centering\includegraphics[scale=1]{clusterer_structure}
	\caption{Iname clusterer structure} \label{clusterer_structure}
\end{figure}

@@

// 5. Return element with regexp
return new Element(
      new ArrayList<String>(),
      rules.get(0).getName(),
      new HashMap<String, Object>(),
      regexp,
      new ArrayList<Attribute>()
      );
\end{verbatim}
The whole submodule structure is in fig. \ref{automaton_merging_state_modules}.

\subsubsection{Data flow}
\jmodule{AutomatonMergingState} works in following steps.
\begin{enumerate}
	\item Create prefix-tree automaton (PTA) from positive examples - right sides of rules.
Call its submodule \jmodule{Au\.to\.ma\.ton\.Simpli\.fier} to modify PTA to some generalized automaton by merging states.
	\item Simplified automaton is then converted to an instance of \code{RegexpAutomaton} (see \ref{section_RegexpAutomaton}) by using clone constructor.
Regexp automaton is automaton with regular expression as symbol on transitions.
In automata theory, such automaton is called extended NFA.
Clone constructing is done by converting each symbol in source automaton to regexp token with that symbol as \code{content}.
	\item \code{RegexpAutomaton} is then passed into second submodule called \jmodule{RegexpAutomatonSimplifier}.
Its job is to derive regular expression from automaton, such that automaton and regular expression represent the same language.
\end{enumerate}
The data flow is in fig. \ref{automaton_merging_state_flow}.

@@

As you can see, it is important that class \code{T} has proper implementation of \code{.equals()} method, as symbols are tested for equality by calling this method on them.
By the way, this is why we cluster all simple data and all sentinels on right sides of rules.	
While adding a concatenation to the automaton, instead of the instances themselves their \emph{representatives} are used. See \ref{section_AutomatonMergingState_structure} for further reference.
This assures that comparisons in automaton are done using Java \code{Object.equals()} reference comparison, but for each cluster member, its representative is only present in automaton.
That results in cluster-equal behaviour.
Example of prefix-tree automaton is in fig. \ref{pta}.

@@

\subsection{StateRemoval (RegexpAutomatonSimplifier)} \label{section_StateRemoval}
\begin{wrapfigure}{R}{40mm}
	\centering\includegraphics[scale=\myscale]{state_removal_modules}
	\caption{Modules of StateRemoval.} \label{state_removal_modules}
	\vspace{10mm}
\end{wrapfigure}
We are using state removal method (see \cite{1224321}) to convert regexp automaton into equivalent regular expression.
This is implemented in \code{StateRemoval} class (on fig. \ref{state_removal_modules}) which implements \code{RegexpAutomatonSimplifier} interface.
Before the algorithm starts, states \emph{superInitial} and \emph{superFinal} are added to automaton, former with $\lambda$-transition to initial state.
From all final states a $\lambda$ transition to \emph{superFinal} state is added.
State removal works by removing states of automaton (and redirecting transitions properly) until there are last to two states - \emph{superInitial} and \emph{superFinal}.
After removing all states, there is only one transition from \emph{superInitial} to \emph{superFinal} state.
That transition has the final regular expression as symbol on it. This one is returned.

@@

Whole TwoStep submodules structure is in fig. \ref{twostep_modules_complete}.
\begin{figure}[tpb]
	\centering\includegraphics[scale=0.37]{twostep_modules_complete}
	\caption{Modules of TwoStep simplifier and their submodules. Filled classes are default selection (best of).} \label{twostep_modules_complete}
\end{figure}

>>>>

%#### INTRO ############################################
\section{Introduction}
In \jmodule{TwoStepSimplifier} we provide the user with a complex solution to grammar simplification problem.
While this is the only simplifier bundled with jInfer, its strong modularity ensures that it can be extended easily. 
For most of the points of extension we indeed provide multiple implementations. 
These vary from trivial proof-of-concept to for example $k,h$-context method for grammar rules inference (see \cite{ahonen}). 
Strong emphasis is put on generic automata, therefore it is easy to plug-in any automaton merging state algorithm.
In this document, we will list all modules and submodules of the simplifier and describe them briefly.

%#### FACTORY ############################################
\section{Factory pattern usage} \label{section_factory}
In \jmodule{TwoStepSimplifier} we employ a factory pattern (as described in \cite[section 4.2]{archdoc}) to divide the module into submodules.
Since service providing classes are being kept as singletons in the NB platform, we use them as factories:
For example:
\begin{verbatim}
@ServiceProvider(service = AutomatonSimplifierFactory.class)
public interface AutomatonSimplifierFactory extends 
                     NamedModule, Capabilities, UserModuleDescription {
  <T> AutomatonSimplifier<T> create();
}
\end{verbatim}
The real submodule interface (called familiarly \emph{worker} interface) is the \code{AutomatonSimplifier<T>} interface (an instance of which is returned by the factory).
For now, it is not important what the worker interface looks like, lets just examine the factory.

In \jmodule{Two\.Step\.Sim\.pli\.fier} we design service providing factory interfaces 
so that they extend \code{Na\.med\.Mo\.du\.le}, \code{Ca\.pa\.bi\.li\.ties} and \code{UserModuleDescription}.
Therefore, the modules must implement following methods:
\begin{verbatim}
String getName();
String getDisplayName();
String getModuleDescription();
List<String> getCapabilities();
String getUserModuleDescription();
\end{verbatim}
Nothing non-standard apart from \code{getUserModuleDescription}.
It returns user-friendly description of the module which is then displayed in properties panels.

The \jmodule{AutomatonSimplifier} works with \code{Automaton} instances, which are generic;
but automaton simplification does not depend on any particular type (of the generic symbol) for automaton transitions, 
so interface \code{AutomatonSimplifier<T>} is generic too.
Factory interface deals with this by defining the \code{create()} method also as generic.

@@

Lets look at \code{AutomatonMergingStateFactory}. This is the factory of a module, which has \jmodule{AutomatonSimplifier} as a submodule.
Its create method (shortened): 
\begin{verbatim}
@Override
public ClusterProcessor<AbstractStructuralNode> create() {
  LOG.debug("Creating new ClusterProcessorAutomatonMergingState.");
  return new AutomatonMergingState(getAutomatonSimplifierFactory(),
                                   getRegexpAutomatonSimplifierFactory());
}
\end{verbatim}
Methods \code{getAutomatonSimplifierFactory} and \code{getRegexpAutomatonSimplifierFactory} are analogical. Here is the former:
\begin{verbatim}
private AutomatonSimplifierFactory getAutomatonSimplifierFactory() {
  Properties p = RunningProject.getActiveProjectProps(getName());

  return ModuleSelectionHelper.lookupImpl(AutomatonSimplifierFactory.class,
            p.getProperty(PROPERTIES_AUTOMATON_SIMPLIFIER,
                          PROPERTIES_AUTOMATON_SIMPLIFIER_DEFAULT));
}
\end{verbatim}
Cluster processor \code{AutomatonMergingState} then receives factories 
of \jmodule{AutomatonSimplifier} and \jmodule{Reg\.exp\.Au\.to\.ma\.ton\.Simplifier} submodules in its constructor.
Cluster processor then may create as many instances of submodule classes as it needs (it may be simplifying more than one automaton).
Thorough this document, we will only mention the worker interface when describing a submodule, 
since all factory interfaces are designed the same way as described previously.

%#### STRUCTURE ############################################
\section{Structure}
\jmodule{TwoStepSimplifier} is implemented in package \code{cz.cuni.mff.ksi.jinfer.twostep}.
Module is divided into two classes: \code{TwoStepSimplifier} (main logic) and \code{TwoStepSimplifierFactory} (lookups, interface to other modules).
The latter is registered as service provider of (and implements the) \code{Simplifier} interface 
(defined in package \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference}).

Its main method \code{start()} receives Initial Grammar in form of
\begin{verbatim}
List<Element> grammar
\end{verbatim}
Each grammar rule is represented by an instance of \code{Element} class. 
Element itself represents the left side of the rule, and its \code{getSubnodes()} method returns 
regular expression representing the right side of the rule (as described in \cite[3.1.1]{archdoc}, they are all concatenations).
Its second parameter is 
\begin{verbatim}
SimplifierCallback callback
\end{verbatim}
Callback which should be invoked when the simplification is done.
On output, simplifier provides Simplified grammar as a parameter of callback function:
\begin{verbatim}
void finished(List<Element> grammar);
\end{verbatim}
Thus simplifier receives Initial Grammar and returns Simplified Grammar, both represented the same way, as a list of elements.

\code{TwoStepSimplifierFactory.start()} creates new \code{TwoStepSimplifier} class instance, providing three factories of submodules to its constructor.
Then it calls \code{simplify(initialGrammar)} method of this instance and its result is passed as parameter to callback function.
And that's all the magic of \code{TwoStepSimplifierFactory} class.

\section{Modular design}
\jmodule{TwoStepSimplifier} is inspired by \cite{1802522} design. Inference proceeds in two steps:
\begin{enumerate}
	\item Clustering the element instances into clusters of (probably) same elements.
	\item Inferring regular expression for each element from examples of element contents taken from all elements in a cluster.
\end{enumerate}
Clustering is delegated to \jmodule{Clusterer} submodule, and the task of regular expression inference 
for each cluster is delegated to \jmodule{ClusterProcessor} submodule. We will examine both of them.
There is also a third submodule called \jmodule{RegularExpressionCleaner}. 
Its purpose is just to beautify output regular expressions, no inference logic is implemented there. 
Modules are drawn on figure \ref{twostep_modules}.
\begin{figure}[tpb]
	\centering\includegraphics[scale=\myscale]{twostep_modules}
	\caption{Submodules of TwoStep simplifier with their implementations} \label{twostep_modules}
\end{figure}
We provide one \jmodule{Clusterer}, four \jmodule{ClusterProcessor} and four \jmodule{RegularExpressionCleaner} implementations.
Each of those will be explained further in this document.

@@

return simplifiedGrammar;
\end{verbatim}
First, it creates a clusterer, gives it all the rules and performs the clustering.
Then, an empty list of elements is created as the simplified grammar.
For each input rule in \code{initialGrammar}, submodule \jmodule{ClusterProcessor} is invoked to infer the regular expression of that element.
Finally, regular expression cleaning is done by a submodule 
and new \code{Element} instance is created as a copy of processed node, but with inferred and cleaned regexp.

\paragraph{Sentinel processing}
There are some specialities in processing sentinel elements (see \cite[3.2]{archdoc}).
Sentinels can be only on the right side of rules. They have to be sent to clusterer, 
as someone further in the chain can ask for cluster of a sentinel element - which appears on right side.
However, sentinels have no content, and so they must not be taken into account when inferring attribute properties 
(whether it is required or optional), since sentinels have no attributes at all.

We will now examine the submodules for clustering, processing and cleaning.

%#### STRUCTURE->Clusterer ############################################
\subsection{Clusterer submodule}
Clustering is implemented in \code{cz.cuni.mff.ksi.jinfer.twostep.clustering} package.
One cluster is represented by the \code{Cluster<T>} class with \code{T} as type parameter of clustered items.
This class simply holds a \code{Set} of members of the cluster, one of which is chosen as a \code{representant}.

Any clusterer has to comply with the \code{Clusterer} interface.
Its purpose is to cluster a bunch of elements (rules) on input into a bunch of \code{Cluster} class instances (clusters) on output.
It has methods \code{add()} and \code{addAll()} for adding the input items to be clustered.
Main part is the method \code{cluster()}, which does the clustering itself.

As this may be a time-consuming operation, method throws \code{InterruptedException}.
Implementation should take care of checking whether thread is user interrupted (see \cite[p. 12]{archdoc}).
To be able to respond to calls to the \code{T getRepresentantForItem(T item)} method, 
every clusterer implementation has to keep all its clusters even after the clustering had finished. 
Using this method, given an item the correct cluster representative for this item has to be returned.

If no such cluster exists (item was not added for clustering before), we recommend throwing an exception rather than returning \code{null}.
Missing an item probably indicates error in the algorithm, rather than normal workflow.
One can get all the clusters (\code{List<Cluster<T>{}>}) from clusterer by calling \code{getClusters()} method.
Basic usage of the clusterer is:
\begin{verbatim}
Clusterer <T> c = new MyContextClusterer<T>();
c.addAll(initialGrammar);
c.cluster();
...
c.getClusters();
or
c.getRepresentantForItem(x);
\end{verbatim}

%#### STRUCTURE->Clusterer->WithAttributes ############################################
\subsubsection{ClustererWithAttributes extended interface}
You might have noticed that the \jmodule{Clusterer} interface and \code{Cluster} class are generic.
They may be used as design pattern not only for clustering elements in the inference process.
To address clustering of elements in more detail, we created \code{Cluster\.er\.With\.At\.tri\.bu\.tes<T, S>} interface, which extends \code{Clusterer<T>} interface.
It adds method \code{List\.<Cluster<S>{}> get\.At\.tri\.bu\.teClusters(T representant)}, implying that each representative of
type \code{T} (that is a representative of some main cluster) has some ``attribute'' clusters associated with it.
Attribute clusters are of type \code{S} and can be retrieved by calling \code{getAttributeClusters(x)}.

@@

%#### STRUCTURE->ClusterProcessor ############################################
\subsection{ClusterProcessor module}
\jmodule{ClusterProcessor} takes rules of one cluster of element and somehow obtains regular expression representing that set of elements. It returns one rule - element with name set to desired name of element in schema (not all elements in a cluster have to have same name, if advanced clustering scheme is used, then processor has to choose the right name for the resulting element) and with subnodes set to regular expression inferred.
It processes attributes of all elements in a cluster to obtain meaningful schema attribute specification. These attributes have to be attached to the resulting element.

Interface of submodule is defined as follows.
\begin{verbatim}
public interface ClusterProcessor<T> {
  T processCluster(
          Clusterer<T> clusterer,
          List<T> rules
          ) throws InterruptedException;
}
\end{verbatim}
Why is the cluster processor given a clusterer instance?
Rules themselves contain information about which elements to process, but clusterer has more information about the topic.
Clusterer can tell you the representative for any arbitrary element from input (not only those elements in rules, but also those that may be on right side of rules). Also, clusterer (if it is with attributes) has information about attributes in each cluster.

@@

%#### STRUCTURE->RegularExpressionCleaner ############################################
\subsection{RegularExpressionCleaner module}
In this last subsection, we examine the \jmodule{RegularExpressionCleaner} interface.
Purpose of this submodule is only to make corrections to regular expression, to make them nicer on output.
For example, it is common that converting automaton to regular expression by state elimination produces nested concatenations such as $(name, (person, id))$.
To convert such expression into $(name, person, id)$, one can implement this interface and connect it to work in chain.
Interface definition is straightforward here.
\begin{verbatim}
public interface RegularExpressionCleaner<T> {
  Regexp<T> cleanRegularExpression(Regexp<T> regexp);
}
\end{verbatim}
Given a regular expression, return cleaned regular expression.
Conversion of regular expression is commonly a recursive task and all our implementations work this way.

@@

%#### DATA FLOW ############################################
\section{Data flow}
\begin{figure}[tpb]
	\centering\includegraphics[scale=1]{twostep_flow}
	\caption{TwoStep data flow} \label{twostep_flow}
\end{figure}
Process illustrated in figure \ref{twostep_flow} is as follows.
\begin{enumerate}
	\item Initial Grammar on input, which is sent into \jmodule{Clusterer}.
	\item Each cluster is sent to submodule \jmodule{ClusterProcessor}, which returns regular expression for that cluster.
	\item Regular expression is sent to \jmodule{RegularExpressionCleaner} submodule for cleaning.
	\item Regular expressions representing all the processed clusters are added to the list of simplified grammar rules.
	\item List of simplified rules is returned (Simplified grammar).
\end{enumerate}

%#### SUBMODULES ############################################
\section{Submodule implementations}
%#### SUBMODULES->Iname ############################################
\subsection{Iname (ClusteterWithAttributes)} \label{section_Iname}
We cluster elements in \code{Iname\.<Abstract\.Structural\.Node, Attri\.bu\.te>} class, which takes \code{Abstract\.Structural\.No\.de} classes to cluster as main, and \code{Attribute} classes as attributes.
Clustering is done based on element's \code{getName()} equality (ignoring case).
For each rule (element), it is first clustered by finding the cluster which representative has the same name - or create new cluster with this element as representative.
Then we process right side of this element rule (that are nodes from \code{getSubnodes()}).
Since right side of rule is always concatenation (\jmodule{TwoStep} accepts only simple grammar representation), we simply take \code{.getSubnodes().getTokens()} list and iterate through it.
Each node on right side, is examined in the following way.
\begin{itemize}
  \item If it is \textbf{simple data}, throw it to \code{SimpleDataClusterer} class which is associated with main element cluster. It does nothing more, than holds all \code{SimpleData} instances in one cluster. But in future it may be replaced to cluster simple data somehow, to obtain meaningful content models in schemas.
  \item If it is \textbf{element} and it is tagged as sentinel - search main clusters to find cluster with representative of same \code{getName()}, or create new cluster with this sentinel as representative. From IGG, sentinels may be only on right sides of rules and since each element in schemas has to be defined, there must exist another element with same name, which is not sentinel (and maybe it will come to process in future). So there can't be cluster with only one sentinel element in it. 
  \item Do nothing otherwise, since it is element, it has to have its subnodes defined, and therefore it is element that is proper grammar rule and therefore it has to be somewhere in Initial Grammar, thus it is already processed or is on schedule.
\end{itemize}
Attributes of element are processed through a helper attribute clusterer, which is created for each element cluster.
We have bunch of elements of same name in a cluster and one attribute clusterer associated with this bunch.
This attribute clusterer is given all attributes instances encountered in all elements that are in bunch.
It clusters them by name, case insensitive.
Each main cluster also has \code{SimpleData} clusterer associated with it.
Whole scheme is on figure \ref{clusterer_structure}.
\begin{figure}[tpb]
	\centering\includegraphics[scale=1]{clusterer_structure}
	\caption{Iname clusterer structure} \label{clusterer_structure}
\end{figure}

@@

// 5. Return element with regexp
return new Element(
      new ArrayList<String>(),
      rules.get(0).getName(),
      new HashMap<String, Object>(),
      regexp,
      new ArrayList<Attribute>()
      );
\end{verbatim}
The whole submodule structure is in figure \ref{automaton_merging_state_modules}.

\subsubsection{Data flow}
\jmodule{AutomatonMergingState} works in following steps.
\begin{enumerate}
	\item Create prefix-tree automaton (PTA) from positive examples - right sides of rules.
Call its submodule \jmodule{Au\.to\.ma\.ton\.Simpli\.fier} to modify PTA to some generalized automaton by merging states.
	\item Simplified automaton is then converted to an instance of \code{RegexpAutomaton} (see \ref{section_RegexpAutomaton}) by using clone constructor.
Regexp automaton is automaton with regular expression as symbol on transitions.
In automata theory, such automaton is called extended NFA.
Clone constructing is done by converting each symbol in source automaton to regexp token with that symbol as \code{content}.
	\item \code{RegexpAutomaton} is then passed into second submodule called \jmodule{RegexpAutomatonSimplifier}.
Its job is to derive regular expression from automaton, such that automaton and regular expression represent the same language.
\end{enumerate}
The data flow is depicted in figure \ref{automaton_merging_state_flow}.

@@

As you can see, it is important that class \code{T} has proper implementation of \code{.equals()} method, as symbols are tested for equality by calling this method on them.
By the way, this is why we cluster all simple data and all sentinels on right sides of rules.	
While adding a concatenation to the automaton, instead of the instances themselves their \emph{representatives} are used. See \ref{section_AutomatonMergingState_structure} for further reference.
This assures that comparisons in automaton are done using Java \code{Object.equals()} reference comparison, but for each cluster member, its representative is only present in automaton.
That results in cluster-equal behaviour.
Example of prefix-tree automaton is in figure \ref{pta}.

@@

\subsection{StateRemoval (RegexpAutomatonSimplifier)} \label{section_StateRemoval}
\begin{wrapfigure}{R}{40mm}
	\centering\includegraphics[scale=\myscale]{state_removal_modules}
	\caption{Modules of StateRemoval.} \label{state_removal_modules}
	\vspace{10mm}
\end{wrapfigure}
We are using state removal method (see \cite{1224321}) to convert regexp automaton into equivalent regular expression.
This is implemented in \code{StateRemoval} class (figure \ref{state_removal_modules}) which implements \code{RegexpAutomatonSimplifier} interface.
Before the algorithm starts, states \emph{superInitial} and \emph{superFinal} are added to automaton, former with $\lambda$-transition to initial state.
From all final states a $\lambda$ transition to \emph{superFinal} state is added.
State removal works by removing states of automaton (and redirecting transitions properly) until there are last to two states - \emph{superInitial} and \emph{superFinal}.
After removing all states, there is only one transition from \emph{superInitial} to \emph{superFinal} state.
That transition has the final regular expression as symbol on it. This one is returned.

@@

Whole TwoStep submodules structure is in figure \ref{twostep_modules_complete}.
\begin{figure}[tpb]
	\centering\includegraphics[scale=0.37]{twostep_modules_complete}
	\caption{Modules of TwoStep simplifier and their submodules. Filled classes are default selection (best of).} \label{twostep_modules_complete}
\end{figure}
